The contact value approximation for the pair distribution function of the inhomogeneous hard-sphere fluid

### 2014-03-26 papers/pair-correlation/paper.tex
make plot a little tighter Tested on bennet. Took 0:17.82.
Rewrite of triplet results. Also edited plots so arrows show on top. Tested on paho-arch. Took 0:54.82.

\begin{figure*}
  \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-contact-3.pdf}
\\
  \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
  \caption{The triplet distribution function
    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
    $\rr_1$ and $\rr_2$ are in contact (a,b) and when $\rr_1$ and
    $\rr_2$ are separated by a distance $2.1\sigma$ (c,d). On the left
    are 2D plots of $g^{(3)}(\rr_1,\rr_2,\rr_3)$ as $\rr_3$
    varies. \plotcomp{$g^{(3)}(\rr_1,\rr_2,\rr_3)$}. Note that an
    asymmetry is introduced in the theoretical results, so their
    values along a mirrored path are also shown, with arrows along the
    curve indicating the path
    direction. \fixme{Explain paths and arrows more.}} \label{fig:triplet-contact-distribution}
\end{figure*}

Figures~\ref{fig:pair-distribution}b and~\ref{fig:pair-distribution}d
show the pair distribution function as plotted along paths illustrated
in Figures~\ref{fig:pair-distribution}a
and~\ref{fig:pair-distribution}c.  These plots compare the CVA and
CVA-S with Monte Carlo results, as well as the approximations of
Sokolowski and Fischer~\cite{sokolowski1992role} and of Fischer and
Methfessel~\cite{fischer1980born} at the same packing fractions of 0.1 and 0.3.
The approach of Fischer and Methfessel is only defined when the two
spheres are in contact, and is therefore only plotted on that segment
of the path.  As an input to both previous approximations and to the
CVA, we use the hard sphere radial distribution function found with
Monte Carlo simulation, interpolated as necessary. We find that both
previous approximations to the pair distribution function predict the
qualitative angular dependence of the pair distribution function at
contact better than this work.  However, in each case the pair
distribution function has a systematic error at contact---either too
high or too low.  While our approximation is smoother than either
of the existing approaches, its errors will have a tendency to cancel
when used in a perturbation expansion.  At higher densities, the
approximation of Fischer and Methfessel requires evaluating the radial
distribution function at densities significantly higher than the
freezing density, which poses numerical difficulties when using the
radial distribution function from simulation.  When the two points
$\rr_1$ and $\rr_2$ are both
more than a radius away from contact, we find that any of these
approaches gives a reasonable prediction, with the CVA-S
underestimating the oscillations in $g_{HS}^{(2)}$, as expected based on
the fit in Fig.~\ref{fig:radial-distribution}.

@@

Figure~\ref{fig:triplet-contact-distribution} compares the triplet
distribution function at a packing fraction of 0.3 computed using the
CVA-S with results from Monte Carlo simulation.  In each subfigure,
$\rr_1$ and $\rr_2$ are held fixed, while the third position $\rr_3$
is varied.  The test-particle position in each case is $\rr_1$, which
is on the left-hand side of the figure.  In the color plots on the
left, the first and second spheres are displayed in gray.  On the
bottom half of figures~\ref{fig:triplet-contact-distribution}a
and~\ref{fig:triplet-contact-distribution}c is the triplet
distribution function as computed using the CVA-S.  The top half of
those figures displays the triplet distribution function from Monte
Carlo simulations.  Figures~\ref{fig:triplet-contact-distribution}b
and~\ref{fig:triplet-contact-distribution}d show the triplet
distribution function along the paths illustrated in the corresponding
2D plots.

Figures~\ref{fig:triplet-contact-distribution}a
and~\ref{fig:triplet-contact-distribution}b show the triplet
distribution function when two of the positions $\rr_1$ and $\rr_2$
are in contact.  The triplet distribution function predicted by the
CVA-S is slightly asymmetric, indicating a failure in thermodynamic
consistency.  Once again, we observe the lack of oscillations at large
distances.  Figures~\ref{fig:triplet-contact-distribution}c
and~\ref{fig:triplet-contact-distribution}d show the triplet
distribution function when two of the positions $\rr_1$ and $\rr_2$
are spaced so that the third could just fit between them.  Again, the
triplet distribution function predicted by the CVA-S is slightly
asymmetric, and lacks the oscillations seen in the Monte Carlo
simulations at large distances.  \fixme{We need to discuss the plots
  and the directions and arrows.} In both cases, our agreement with
simulation is comparable to that of previous approximations.

>>>>

\begin{figure*}
  \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-contact-3.pdf}
  \vspace{-0.7cm}
\\
  \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
  \vspace{-0.7cm}
  \caption{The triplet distribution function
    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
    $\rr_1$ and $\rr_2$ are in contact (a,b) and when $\rr_1$ and
    $\rr_2$ are separated by a distance $2.1\sigma$ (c,d). On the left
    are 2D plots of $g^{(3)}(\rr_1,\rr_2,\rr_3)$ as $\rr_3$
    varies. \plotcomp{$g^{(3)}(\rr_1,\rr_2,\rr_3)$}. Note that an
    asymmetry is introduced in the theoretical results, so their
    values along a mirrored path are also shown, with arrows along the
    curve indicating the path direction, as explained in the
    text.} \label{fig:triplet-contact-distribution}
\end{figure*}

Figures~\ref{fig:pair-distribution}b and~\ref{fig:pair-distribution}d
show the pair distribution function as plotted along paths illustrated
in Figures~\ref{fig:pair-distribution}a
and~\ref{fig:pair-distribution}c.  These plots compare the CVA and
CVA-S with Monte Carlo results, as well as the approximations of
Sokolowski and Fischer~\cite{sokolowski1992role} and of Fischer and
Methfessel~\cite{fischer1980born} at the same packing fractions of 0.1 and 0.3.
The approach of Fischer and Methfessel is only defined when the two
spheres are in contact, and is therefore only plotted on that segment
of the path.  As an input to both previous approximations and to the
CVA, we use the hard sphere radial distribution function found with
Monte Carlo simulation, interpolated as necessary. We find that both
previous approximations to the pair distribution function predict the
qualitative angular dependence of the pair distribution function at
contact better than this work.  However, in each case the pair
distribution function has a systematic error at contact---either too
high or too low.  While our approximation is smoother than either
of the existing approaches, its errors will have a tendency to cancel
when used in a perturbation expansion.  At higher densities, the
approximation of Fischer and Methfessel requires evaluating the radial
distribution function at densities significantly higher than the
freezing density, which poses numerical difficulties when using the
radial distribution function from simulation.  When the two points
$\rr_1$ and $\rr_2$ are both
more than a radius away from contact, we find that any of these
approaches gives a reasonable prediction, with the CVA-S
underestimating the oscillations in $g^{(2)}$, as expected based on
the fit in Fig.~\ref{fig:radial-distribution}.

@@

Figures~\ref{fig:triplet-contact-distribution}a
and~\ref{fig:triplet-contact-distribution}c compare the triplet
distribution function at a packing fraction of 0.3 computed using the
CVA-S with results from Monte Carlo simulations. In
Figure~\ref{fig:triplet-contact-distribution}a the spheres at $\rr_1$
and $\rr_2$ are in contact, in
Figure~\ref{fig:triplet-contact-distribution}c they are spaced so that
a third sphere can just fit between them, and in both figures $\rr_3$
is varied. The test-particle position for the CVA-S in each case is
$\rr_1$, which is on the left-hand side of the figure. We see
reasonable agreement, with flatter oscillations in the CVA-S at larger
distances. Also, the Monte Carlo results have perfect left-right
symmetry, while the CVA-S has an asymmetry introduced with
the test particle due to imperfections in the pair distribution
function. Most notably, the CVA-S has less angular dependence at
contact with the test particle than it does with the sphere at
$\rr_2$.

Figures~\ref{fig:triplet-contact-distribution}b
and~\ref{fig:triplet-contact-distribution}d show the triplet
distribution function as plotted along the paths illustrated in
Figures~\ref{fig:triplet-contact-distribution}a
and~\ref{fig:triplet-contact-distribution}c. As the theoretical
results are asymmetric, they are also plotted along similar paths that
travel to the left, along the test particle. The different paths are
indicated by arrows along the curves, with right-facing arrows
indicating the paths shown in
Figures~\ref{fig:triplet-contact-distribution}a and
\ref{fig:triplet-contact-distribution}c, and left-facing arrows
indicating paths traveling along the test particle.  As the work of
Fischer and Methfessel is only defined in contact, it is only plotted
in Figure~\ref{fig:triplet-contact-distribution}b and only along the
portion of the path shown that is in contact with $\rr_2$, and arrows
are omitted. Again, the CVA-S underestimates the oscillations,
especially in contact with the test particle when the points $\rr_1$
and $\rr_2$ are separated by a distance of $2.1\sigma$, as seen in
Figure~\ref{fig:triplet-contact-distribution}d. Otherwise, all methods
perform similarly, with the method of Fischer and Methfessel giving
the best results when the spheres at $\rr_1$ and $\rr_2$ are in
contact.

% Figure~\ref{fig:triplet-contact-distribution} compares the triplet
% distribution function at a packing fraction of 0.3 computed using the
% CVA-S with results from Monte Carlo simulation.  In each subfigure,
% $\rr_1$ and $\rr_2$ are held fixed, while the third position $\rr_3$
% is varied.  The test-particle position in each case is $\rr_1$, which
% is on the left-hand side of the figure.  In the color plots on the
% left, the first and second spheres are displayed in gray.  On the
% bottom half of figures~\ref{fig:triplet-contact-distribution}a
% and~\ref{fig:triplet-contact-distribution}c is the triplet
% distribution function as computed using the CVA-S.  The top half of
% those figures displays the triplet distribution function from Monte
% Carlo simulations.  Figures~\ref{fig:triplet-contact-distribution}b
% and~\ref{fig:triplet-contact-distribution}d show the triplet
% distribution function along the paths illustrated in the corresponding
% 2D plots.

% Figures~\ref{fig:triplet-contact-distribution}a
% and~\ref{fig:triplet-contact-distribution}b show the triplet
% distribution function when two of the positions $\rr_1$ and $\rr_2$
% are in contact.  The triplet distribution function predicted by the
% CVA-S is slightly asymmetric, indicating a failure in thermodynamic
% consistency.  Once again, we observe the lack of oscillations at large
% distances.  Figures~\ref{fig:triplet-contact-distribution}c
% and~\ref{fig:triplet-contact-distribution}d show the triplet
% distribution function when two of the positions $\rr_1$ and $\rr_2$
% are spaced so that the third could just fit between them.  Again, the
% triplet distribution function predicted by the CVA-S is slightly
% asymmetric, and lacks the oscillations seen in the Monte Carlo
% simulations at large distances.  \fixme{We need to discuss the plots
%   and the directions and arrows.} In both cases, our agreement with
% simulation is comparable to that of previous approximations.
